---
title: "STAT 479 HW1"
output: rmarkdown::pdf_document
author: "Madison Polley"
header-includes:
    - \usepackage {hyperref}
    - \hypersetup {colorlinks = true, linkcolor = red, urlcolor = red}
---

```{r, echo = FALSE}
library("knitr")
opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE)
```

## Instructions

1. Submit your solutions to the exercises below before **February 5 at 11:59pm
CST**.
2. Prepare your solutions as Rmarkdown documents. For problems in vega-lite,
copy your code into the `.Rmd`, making sure to enclose them within `js` blocks,

`````
````js
// your code here
````
`````

so that the syntax is properly highlighted, e.g.,

````js
// your code here
````

3. For vega-lite problems, include a screenshot of your result.
4. We give example figures below to guide your work. However, view these only as
suggestions -- we will keep an eye out for improvements over our plots.
5. Include two files in your submission, (a) a pdf of the compiled `.Rmd` file
and (b) the original `.Rmd` file.

## Rubric

2 problems below will be graded according to,

* Correctness: For plots, the displays meet the required specifications. For
conceptual questions, all parts are accurately addressed.
* Attention to detail: Writing is clear and designs are elegant. For example, no
superfluous marks are included, all axes are labeled, text is neither too small
nor too large.
* Code quality (if applicable): Code is concise but readable, properly formatted
and commented.

The remaining problems will be graded for completeness.

## Problems

### (1) Ikea Furniture

The dataset below shows prices of pieces of Ikea furniture. We will compare
prices of different furniture categories and label the (relatively few) articles
which cannot be bought online.

```{r}
library("readr")
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggridges)
library(scales)
ikea <- read_csv("https://uwmadison.box.com/shared/static/iat31h1wjg7abhd2889cput7k264bdzd.csv")
```
a. Make a plot that shows the relationship between the `category` of furniture
and `price` (on a log-scale). Show each `item_id` as a point -- do not aggregate
to boxplots or ridgelines -- but make sure to jitter and adjust the size the
points to reduce the amount of overlap. _Hint: use the `geom_jitter` layer._

```{r}
ggplot(ikea, aes(price, category)) + geom_jitter( size = 0.7) + scale_x_log10()
```


b. Modify the plot in (a) so that categories are sorted from those with highest
to lowest average prices.

```{r}
ikea$category = as.factor(ikea$category)
ikea %>%
  mutate(class = reorder(category, price, .fun='mean')) %>%
  ggplot(aes(x=price, y = reorder(category, price))) + geom_jitter( size = 0.7) + scale_x_log10() + labs(y = "Category")
```


c. Color points according to whether they can be purchased online. If they
cannot be purchased online, add a text label giving the name of that item of
furniture. An example result is given by Figure 1.

```{r}
unavailable = ikea %>% filter(sellable_online == FALSE)

ikea %>%
  mutate(class = reorder(category, price, .fun='mean')) %>%
  ggplot(aes(x=price, y = reorder(category, price))) + geom_jitter( aes(color = sellable_online), size = 0.7) + scale_x_log10() + labs(y = "Category") + geom_text_repel(data = unavailable, label = unavailable$name, size = 3, line = 3)
```


### (2) Penguins

The data below measures properties of various antarctic penguins.

```{r}
penguins <- read_csv("https://uwmadison.box.com/shared/static/ijh7iipc9ect1jf0z8qa2n3j7dgem1gh.csv")
```
Using either vega-lite or ggplot2, create a single plot that makes it easy to
answer both of these questions,

```{r}
ggplot(penguins, aes(bill_length_mm, species)) + geom_jitter(aes(color = bill_depth_mm, shape = island)) + scale_color_gradientn(colors = rainbow(5)) + labs(x = "Bill Length in Millimeters")
```


i) How is bill length related to bill depth within and across species? 

**It appears that for all penguins, the more shallow the bill, the the shorter the length. Looking at all penguins as a whole, it appears that short and long bill lengths yield deep bills, but average bill length produces a more shallow bill.**

ii) On which islands are which species found?

**It appears that Gentoo penguins are found on Biscoe Island, Chinstrap penguins on Dream Island, and Adelle penguins are on all three islands.**

(Notice that the answer to part (i) is an example of Simpson's paradox!)


### (3) 2012 London Olympics

This exercise is similar to the Ikea furniture one, except that it will be
interactive. The data at this
[link](https://uwmadison.box.com/s/rzw8h2x6dp5693gdbpgxaf2koqijo12l) describes
all participants in the London 2012 Olympics. From an observable notebook, the
following code can be used to derive a new variable with a jittered Age
variable, which will be useful in part (a).

````js
import { vl } from "@vega/vega-lite-api"
import { aq, op } from "@uwdata/arquero"
data_raw = aq.fromCSV(await FileAttachment("All London 2012 athletes - ALL ATHLETES.csv").text())
data = data_raw.derive({Age_: d => d.Age + 0.25 * Math.random() })
````

a. Create a layered display that shows (i) the ages of athletes across sports
and (ii) the average age within each sport. Use different marks for participants
and for averages. To avoid overplotting, use the jittered `Age_` variable
defined in the code block above.

````js
{
  const all = vl.markPoint().data(data).encode(
    vl.y().fieldN("Sport").title("Sport"),
    vl.x().fieldQ("Age_").title("Age")
  );
  const point = vl.markSquare({color: 'orange'}).data(data).encode(
    vl.y().fieldN("Sport"),
    vl.x().fieldQ("Age_").average("Age_"),
  );
  
  return vl.layer(all, point).render();
}
````

b. Sort the sports from lowest to highest average age. Add a tooltip so that
hovering over an athlete shows their name. Your results should look something
like the display in Figure \ref{fig:3}.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Athletes.png}
  \caption{Vega-lite result for Problem (3). Hovering over a tick mark
  shows the name of the athlete.}
  \label{fig:3}
\end{figure}

````js
{
  const all = vl.markPoint().data(data).encode(
    vl.y().fieldN("Sport").sort(vl.average("Age_").order('ascending')),
    vl.x().fieldQ("Age_").title("Age"),
    vl.tooltip().fieldN("Name")
  );
  const point = vl.markSquare({color: "orange"}).data(data).encode(
    vl.y().fieldN("Sport").sort(vl.average("Age_").order('ascending')),
    vl.x().fieldQ("Age_").average("Age_"),
  );
  
  return vl.layer(all, point).render();
}
````

### (4) Traffic

In lecture, we looked at the `geom_density_ridges` function. In this exercise,
we will instead use `geom_ridgeline`, which is useful whenever the heights of the
ridges have been computed in advance. We will use the traffic data read in
below.

```{r}
traffic <- read_csv("https://uwmadison.box.com/shared/static/x0mp3rhhic78vufsxtgrwencchmghbdf.csv")
```

Each row is a timepoint of traffic within a city in Germany. Using
`geom_ridges`, make a plot of traffic over time, within each of the cities. An
example result is shown below.

```{r}
traffic$name = as.factor(traffic$name)
ggplot(traffic, aes(date, y = reorder(name, value, mean), height = value)) + geom_ridgeline()
```


### (5) Language Learning

This problem will look at a simplified version of the data from the study _A
critical period for second language acquisition: Evidence from 2/3 million
English speakers_, which measured the effect of the the age of initial language
learning on performance in grammar quizzes. We have downloaded the raw data from
the supplementary material and reduced it down to the average and standard
deviations of test scores within (initial learning age) $\times$ (current
age-group) combinations. We have kept a column `n` showing how many participants
were used to compute the associated statistics. The resulting data are available
[here](https://uwmadison.box.com/shared/static/m53dea9w5ipczs3d7nqdnxqxx5ao501b.csv).

a) Using the `.derive()` command in arquero, create two new fields, `low` and
`high`, giving confidence intervals for the means in each row. That is, derive
new variables according to $\hat{x} \pm 2 * \frac{1}{\sqrt{n}}\hat{\sigma}$.

````js
data2 = data_raw2.derive({low: d => (d.avg_correct - (2 * (1/Math.sqrt(d.n)) * d.sd_correct)), high: d => (d.avg_correct + (2 * (1/Math.sqrt(d.n)) * d.sd_correct))})
````

b) Create a `markArea`-based ribbon plot showing confidence intervals for
average test scores as a function of starting age. Include a line for the
average score within that combination. An example result is shown in Figure
\ref{fig:5}. Interpret the results of the study.

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{Language.png}
  \caption{Example result for problem (5).}
  \label{fig:5}
\end{figure}

````js
{
  const confint = vl.markArea({opacity: 0.3})
    .data(data2)
    .encode(
      vl.x().fieldQ("Eng_start").title("Age They Began English"),
      vl.y().fieldQ('high').scale({domain: [0.75, 1]}).title("Average Score with Confidence Band"),
      vl.y2().fieldQ('low'),
      vl.color().fieldN('age_group')
    );

  const line = vl.markLine()
    .data(data2)
    .encode(
      vl.x().fieldQ("Eng_start"),
      vl.y().fieldQ("avg_correct").scale({domain: [0.75, 1]}),
      vl.color().fieldN('age_group')
    );

  return vl.layer(confint, line).render();
}
````

### (6) Deconstruction

Take a static screenshot from any of the visualizations in this
[article](https://www.theguardian.com/us-news/ng-interactive/2017/dec/20/bussed-out-america-moves-homeless-people-country-study),
and deconstruct its associated visual encodings.

**I looked at the blue map of homeless rates per 100,000.**

a) What do you think was the underlying data behind the current view? What where
the rows, and what were the columns?

**The underlying data was of the homeless rate per 100,000 in each state in the US. The rows are the individual states. The only column is the homeless rate per 100,000 unless the data also includes the latitude and longitude in order to plot the states where they are.**

b) What were the data types of each of the columns?

**The data type of the homeless rate per 100,000 may be an ordered categorical variable since it only gives 4 categories in the colors. The latitude and longitude are numeric.**

c) What encodings were used? How are properties of marks on the page derived
from the underlying abstract data?

**It appears that they used vl.x, vl.y, vl.color, and probably something to make their marks in the shapes of the states. The colors of the points on the page are from the categorical variable for the homeless rate. The location of the states could be from latitude and longitude. The shape would be from the state.**

c) Is multi-view composition being used? If so, how?

**No.**
